---
title: "Data Visualization"
author: "Erica Dale"
date: "2022-08-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, #shows code
                      message = FALSE,
                      warning = FALSE)
library(tidyverse)
library(here)
library(janitor)
library(patchwork)
library(ggrepel)
library(gghighlight)
library(paletteer)
library(ggExtra)
library(ggbeeswarm)
library(gapminder)
library(sf)
```

### Read in Lizard data
```{r}
lizards <- read_csv(here("data_tidy", "lizards.csv"))
```

Refresher
Plot total length vs weight
```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point()
```

Create a histogram
```{r}
ggplot(data = lizards, aes(x = total_length)) +
  geom_histogram(color = "orange",
                 fill = "purple",
                 size = 1,
                 linetype = "dotted") #inside aes adds legend
```

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(
    shape = 23,
    color = "cyan4",
    fill = "purple",
    size = 4,
    alpha = 0.4
  )
```

### Changing aesthetic by variable
```{r}
ggplot(data = lizards, aes(x= total_length, 
                           y = weight)) +
  geom_point(aes(color = common_name, # lisa frank -esque
                 size = total_length),
             alpha = 0.5) +
  theme_minimal()
```

### Faceting
```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  facet_wrap(~common_name, scales = "free")
```


### Reordering

#### Summarize observation counts
```{r}
lizcount <- lizards |> 
  group_by(common_name) |> 
  summarize(tot_lizards = n())
lizcount

lizcount <- lizards |> 
  count(common_name)
lizcount
```

```{r}
ggplot(lizcount, aes(y = fct_reorder(common_name, n), 
                     x = n)) +
  geom_col()
```

```{r}
ggplot(lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name,
                 shape = common_name) +
  facet_wrap(~common_name, scales = "free_y") +
  theme_dark() # +
#  labs(x = "Total Length (mm)",
#       y = "Weight (grams)",
#       title = "Jornada Basin lizard sizes",
#       subtitle = "All lizards follow standars length-weight relationship, with Western Whiptails being largest.",
#       caption = "Data: Jornada Basin LTER, CCBY")

```


### Fun with scales

#### Change gradient colors
```{r}
ggplot(lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) +
  scale_color_gradient(low = "purple", high = "orange")   # changing the gradient
```

#### Specify gradient range
```{r}
ggplot(lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) +
  scale_color_gradientn(colors = c("magenta", "cyan4", "darkorchid3", "dodgerblue")) # add a range for colors
```

#### Color range in bins
```{r}
ggplot(lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) +
  scale_color_steps(low = "red", high = "black") # Sets colors into bins
```

#### Diverging color scheme
```{r}
ggplot(lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = total_length)) +
  scale_color_steps2(low = "purple",
                     mid = "white",
                     high = "orange",
                     midpoint = 150,
                     breaks = c(50, 75, 150, 180, 220, 280)) # Diverging color scheme
```

#### Manually set color sequences
```{r}
ggplot(lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = total_length)) +
  scale_color_stepsn(colors = c("orange", "red", "purple"),
                     breaks = seq(from = 0, 
                                  to = 300, 
                                  by = 60)) # Create equal length
```

### Pallets

#### Discrete color pallets
```{r}
glimpse(palettes_d_names)
```

#### Reorder factors
```{r}
ggplot(lizards, aes(y = common_name, x = total_length)) +
  geom_boxplot(aes(fill = common_name))

lizards_med <- lizards |> 
  mutate(common_name = fct_reorder(common_name, total_length, .fun = median)) #reorder common_name by median of total_length

ggplot(lizards_med, aes(y = common_name, x = total_length)) +
  geom_boxplot(aes(fill = common_name), show.legend = FALSE) +
  scale_fill_paletteer_d(palette = "ggsci::planetexpress_futurama") + #from what package, and what pallet name
  theme_minimal()
```

#### Themes: Make a terrible graph
##### Add custom annotation
```{r}
ggplot(lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) +
  theme(panel.grid.major.x = element_line(color = "red"), #hyper customization
        panel.grid.minor.y = element_blank(),
        axis.title.x = element_text(color = "purple", size = 10),
        axis.text.y = element_text(color = "green"),
        text = element_text(size = 18),
        panel.background = element_rect(color = "purple", fill = "violet")) +
  annotate("text", x = 300, y = 50, label = "Wooo", color = "blue") + # type of addition, where, what
  geom_vline(xintercept = 250, linetype = "dashed", color = "orange", size = 4)
  
```

### `ggrepel` Labels
```{r}
wws_lizards <- lizards |> 
  filter(common_name == "western whiptail",
         site == "sand")

ggplot(wws_lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  geom_text(aes(label = toe_num)) # built in: obnoxious, auto centers label over point

ggplot(wws_lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  geom_text_repel(aes(label = toe_num))
```

```{r}
gapminder |> 
  filter(year == 2002,
         continent == "Europe") |> 
  ggplot(aes(x = gdpPercap, y = lifeExp)) +
  geom_point(aes(color = country), show.legend = FALSE) +
  geom_text_repel(aes(label = country), size = 3)

```

### Highlight

```{r}
ggplot(lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  gghighlight(weight > 30, label_key = toe_num)
```

### Map
```{r}
jornada_veg <- read_sf(here("data_raw", "spatial_vegetation", "doc.kml"))

ggplot() +
  geom_sf(data = jornada_veg,
          aes(fill = Name),
          color = NA) +  # removes outlines
  theme_minimal() +
  scale_fill_paletteer_d(palette = "ggthemes::manyeys")
```

### Heatmap

```{r}
lizard_counts <- lizards |> 
  mutate(date = lubridate::mdy(date)) |> ## Updating date column
  count(year = lubridate::year(date), # Count by year and common name
        common_name) |> 
  drop_na() 

ggplot(lizard_counts, aes(x = year, y = common_name)) +
  geom_tile(aes(fill = n), show.legend = FALSE) +
  geom_text(aes(label = n), color = "white", size = 3) +
  scale_fill_gradientn(colors = c("navy", "red", "orange")) +
  theme_minimal()
```

### Beeswarm Plot
good for confined number of observations

```{r}
whiptails <- lizards |> 
  filter(common_name == "western whiptail") |> 
  drop_na(total_length, weight) # Drops only if total weight or length is NA

whip_b_plot <- ggplot(data = whiptails, aes(x = sex, y = weight)) +
  geom_beeswarm(color = "purple") +
  geom_boxplot(fill = NA)
```

### Marginal Plot

```{r}
whiptail_plot <- ggplot(data = whiptails, aes(x = total_length, y = weight)) +
  geom_point(aes(color = sex), shape = 6) +
  theme(legend.position = "bottom")

ggMarginal(whiptail_plot, type = "boxplot", groupColour = TRUE)
```


### Patchwork

```{r}
((whip_b_plot + whiptail_plot) / whiptail_plot) + whip_b_plot & theme_dark()

```

